### 2021-09-03 manuscript/methodology.tex
until structure of Axx, Ayy and Azz matrix components

The computational cost required for estimating the classical solution can be very high
when dealing with large datasets. In the following subsections, we will show how to 
explore the structure of the sensitivity matrix $\mathbf{A}$ and use this to
efficiently solve the least-squares normal equations (equation \ref{eq:normal-equations}).

@@

To access the structure of the sensitivity matrix $\mathbf{A}$ 
(equation \ref{eq:predicted-data-vector_mag}), let us first rewrite its elements 
$a_{ij}$ (equation \ref{eq:aij_mag}) in the following way:
\begin{equation}
\begin{split}
a_{ij} = a^{xx}_{ij} + a^{xy}_{ij} + a^{xz}_{ij} + a^{yy}_{ij} + a^{yz}_{ij} + a^{zz}_{ij} \: ,
\end{split}
\label{eq:aij_mag_expand}
\end{equation}
where
\begin{equation}
a^{\alpha\beta}_{ij} = c_{m} \, \frac{\mu_{0}}{4\pi} 
\left( F_{\alpha} h_{\beta} + F_{\beta} h_{\alpha} \right) h^{\alpha\beta}_{ij}
\: , \quad \alpha, \beta = x, y, z \: ,
\label{eq:aij_alpha_beta}
\end{equation}
are defined by the elements of $\hat{\mathbf{F}}$ 
(equation \ref{eq:unit_vector_F}), $\hat{\mathbf{u}}$ (equation \ref{eq:u_hat}) and 
$\mathbf{H}_{ij}$ (equations \ref{eq:Hij} and \ref{eq:hij_alpha_beta}).
Then, we can rewrite the sensitivity matrix $\mathbf{A}$ 
(equation \ref{eq:predicted-data-vector_mag}) according to:
\begin{equation}
\mathbf{A} = \mathbf{A_{xx}} + \mathbf{A_{xy}} + \mathbf{A_{xz}} + 
\mathbf{A_{yy}} + \mathbf{A_{yz}} + \mathbf{A_{zz}} \: ,
\label{eq:A_expand}
\end{equation}
where $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ are $N \times M$ matrices with elements 
$ij$ defined by $a^{\alpha\beta}_{ij}$ (equation \ref{eq:aij_alpha_beta}).

@@

By using equations \ref{eq:xi}--\ref{eq:l-y-oriented} to define the coordinates $x_{i}$ and 
$y_{i}$ of the observation points and $x_{j}$ and $y_{j}$ of the equivalent sources, we can
rewrite the elements $h^{\alpha\beta}_{ij}$ (equation \ref{eq:hij_alpha_beta}) of matrix 
$\mathbf{H}_{ij}$ (equation \ref{eq:Hij}) as follows:
\begin{equation}
h^{xx}_{ij} = 
\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
\label{eq:hxx_regular}
\end{equation}
\begin{equation}
h^{yy}_{ij} = 
\frac{3 \left( \Delta l_{ij} \, \Delta_{y} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
\label{eq:hyy_regular}
\end{equation}
\begin{equation}
h^{zz}_{ij} = 
\frac{3 \Delta_{z}^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
\label{eq:hzz_regular}
\end{equation}
\begin{equation}
h^{xy}_{ij} = 
\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right)\left( \Delta l_{ij} \, \Delta_{y} \right)}{r_{ij}^{5}} \: ,
\label{eq:hxy_regular}
\end{equation}
\begin{equation}
h^{xz}_{ij} = 
\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right) \Delta_{z}}{r_{ij}^{5}}
\label{eq:hxz_regular}
\end{equation}
and
\begin{equation}
h^{yz}_{ij} = 
\frac{3 \left( \Delta l_{ij} \, \Delta_{y} \right) \Delta_{z}}{r_{ij}^{5}}
\label{eq:hyz_regular}
\end{equation}
where $\Delta_{z} = z_{c} - z_{0}$, $\Delta k_{ij} = k(i) - k(j)$ (equations \ref{eq:k-x-oriented} or \ref{eq:k-y-oriented}), $\Delta l_{ij} = l(i) - l(j)$ (equations \ref{eq:l-x-oriented} or \ref{eq:l-y-oriented}) and
\begin{equation}
\frac{1}{r_{ij}} = 
\frac{1}{\sqrt{\left( \Delta k_{ij} \, \Delta_{x} \right)^{2} + \left( \Delta l_{ij} \, \Delta_{y} \right)^{2} + \Delta_{z}^{2}}} \: .
\label{eq:1_rij_regular}
\end{equation}

@@

By using equations \ref{eq:hxx_regular}--\ref{eq:1_rij_regular} to compute 
$a^{\alpha\beta}_{ij}$ (equation \ref{eq:aij_alpha_beta}), we obtain well-structured 
block matrices $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}).
All of them are formed by a grid of $Q \times Q$, where each block has $P \times P$ 
elements. The blocks disposed along the same block diagonal are equal to each other.
In each block, the elements disposed on the same diagonal are also equal to each other.
Such matrices are called Block-Toeplitz Toeplitz-Block (BTTB) and they can be conveniently
represented by using \textit{block indices} $q$ and $p$ \citep{takahashi2020convolutional}.
Here, each block is represented by $\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$, 
$q = -Q+1,...,0,..., Q-1$, and their elements by $a^{\alpha\beta}_{qp}$, 
$p = -P+1,...,0,...,P-1$. 
Positive (negative) indices $q$ represent blocks $\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$ 
located below (above) the main block diagonal of $\mathbf{A}_{\boldsymbol{\alpha\beta}}$.
Similarly, positive (negative) indices $p$ represent elements $a^{\alpha\beta}_{qp}$ 
located below (above) the main diagonal of $\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$.
According to this notation, $q$ defines the block $\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$
and $p$ the diagonal where the element $a^{\alpha\beta}_{qp}$ is located within this block, so that
matrices $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ 
can be written, in a general form, as a grid of $Q \times Q$ blocks
\begin{equation}
\mathbf{A}_{\boldsymbol{\alpha\beta}} = \begin{bmatrix}
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{0}   & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{-1} & \cdots          & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{-Q+1} \\
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{1}   & \ddots          & \ddots          & \vdots           \\ 
\vdots           & \ddots          & \ddots          & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{-1}   \\
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{Q-1} & \cdots          & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{1}  & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{0}
\end{bmatrix}_{N \times N} \: ,
\label{eq:BTTB_A_alpha_beta}
\end{equation}
with blocks $\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$ given by
\begin{equation}
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q} = \begin{bmatrix}
a^{\alpha\beta}_{q0}   & a^{\alpha\beta}_{q(-1)} & \cdots  & a^{\alpha\beta}_{q(-P+1)} \\
a^{\alpha\beta}_{q1}   & \ddots     & \ddots  & \vdots       \\ 
\vdots      & \ddots     & \ddots  & a^{\alpha\beta}_{q(-1)}   \\
a^{\alpha\beta}_{q(P-1)} & \cdots     & a^{\alpha\beta}_{q1}  & a^{\alpha\beta}_{q0}
\end{bmatrix}_{P \times P} \: .
\label{eq:Aq_block}
\end{equation}
It is important to stress that $a^{\alpha\beta}_{qp}$ (equation \ref{eq:Aq_block}), 
$q = -Q+1,...,0,..., Q-1$, $p = -P+1,...,0,...,P-1$, represents an alternative notation
to describe the elements $a^{\alpha\beta}_{ij}$ (equation \ref{eq:aij_alpha_beta}) of the matrix
components $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}) in terms of 
the block indices $q$ and $p$. This notation allows us to verify that: (i) different combinations
of indices $i$ and $j$ result in elements $a^{\alpha\beta}_{ij}$ having the same numerical 
value; (ii) these elements are located on the diagonal $p$ of a block 
$\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$ (equation \ref{eq:Aq_block}); (iii) the same block 
$\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$ appears along the block diagonal $q$ of the 
matrix component $\mathbf{A}_{\boldsymbol{\alpha\beta}}$ (equation \ref{eq:BTTB_A_alpha_beta}).

The relationship between a given element $a^{\alpha\beta}_{ij}$ (equation \ref{eq:aij_alpha_beta})
and an element $a^{\alpha\beta}_{qp}$ is defined by the block indices $q$ and $p$ 
and the integer functions $\Delta k_{ij}$ and $\Delta l_{ij}$ 
(equations \ref{eq:hxx_regular}--\ref{eq:1_rij_regular}).
For $x$-\textit{oriented grids}, $Q = N_{y}$, $P = N_{x}$ and the block indices
$q$ and $p$ are given by:
\begin{equation}
q(i, j) = \Delta l_{ij} = l(i) - l(j)
\label{eq:q-x-oriented}
\end{equation}
and
\begin{equation}
p(i, j) = \Delta k_{ij} =  k(i) - k(j) \: ,
\label{eq:p-x-oriented}
\end{equation}
where $k(i)$ and $k(j)$ are given by equation \ref{eq:k-x-oriented} and 
$l(i)$ and $l(j)$ by equation \ref{eq:l-x-oriented}.
For $y$-\textit{oriented grids}, $Q = N_{x}$, $P = N_{y}$ and the block indices
$q$ and $p$ are given by:
\begin{equation}
q(i, j) = \Delta k_{ij} = k(i) - k(j) 
\label{eq:q-y-oriented}
\end{equation}
and
\begin{equation}
p(i, j) = \Delta l_{ij} = l(i) - l(j) \: ,
\label{eq:p-y-oriented}
\end{equation}
where $k(i)$ and $k(j)$ are given by equation \ref{eq:k-y-oriented} and 
$l(i)$ and $l(j)$ by equation \ref{eq:l-y-oriented}.
Equations \ref{eq:q-x-oriented}--\ref{eq:p-y-oriented} differ from those presented by
\citet{takahashi2020convolutional} in the absence of the module.

@@

@@

%======================================================================================
%\subsection{Structure of matrices components $\mathbf{H_{xx}}$, $\mathbf{H_{yy}}$ and $\mathbf{H_{zz}}$}
%======================================================================================

The principal components $\mathbf{H_{xx}}$, $\mathbf{H_{yy}}$ and $\mathbf{H_{zz}}$ (equations \ref{eq:hxx_mag}, \ref{eq:hyy_mag} and \ref{eq:hzz_mag}, respectively) of matrix $\mathbf{H}$ (equation \ref{eq:Hi}) are symmetric-Block-Toeplitz symmetric-Toeplitz-Block matrices. This means that $\mathbf{H_{xx}}$, $\mathbf{H_{yy}}$ and $\mathbf{H_{zz}}$ are Toeplitz and symmetric by its blocks and each of the blocks are symmetric Toeplitz matrices. 
For example, $\mathbf{H_{xx}}$ can be described by the \textit{block index} $q$ that represents the block diagonals of this matrix as a grid of $Q \times Q$ blocks $\mathbf{H}^{q}_\mathbf{xx}$, $q = 0, \dots, Q - 1$:
\begin{equation}
\mathbf{H_{xx}} = \begin{bmatrix}
\mathbf{H}^{0}_\mathbf{xx}  & \mathbf{H}^{1}_\mathbf{xx} & \cdots         & \mathbf{H}^{Q-1}_\mathbf{xx} \\
\mathbf{H}^{1}_\mathbf{xx}  & \mathbf{H}^{0}_\mathbf{xx} & \ddots         & \vdots           \\ 
\vdots           & \ddots         & \ddots         & \mathbf{H}^{1}_\mathbf{xx}   \\
\mathbf{H}^{Q-1}_\mathbf{xx} & \cdots         & \mathbf{H}^{1}_\mathbf{xx} & \mathbf{H}^{0}_\mathbf{xx}                
\end{bmatrix}_{N \times N} \: .
\label{eq:BTTB_Hxx}
\end{equation}
And each diagonal of the blocks are represented by $P \times P$ elements $h^{q}_{p}$, $p = 0, \dots, P - 1$:
\begin{equation}
\mathbf{H}^{q}_\mathbf{xx} = \{h^{q}_p\} = \begin{bmatrix}
h^{q}_{0}   & h^{q}_{1} & \cdots    & h^{q}_{P-1} \\
h^{q}_{1}   & h^{q}_{0} & \ddots    & \vdots           \\ 
\vdots      & \ddots    & \ddots    & h^{q}_{1}   \\
h^{q}_{P-1} & \cdots    & h^{q}_{1} & h^{q}_{0}                 
\end{bmatrix}_{P \times P} \: .
\label{eq:Hxx_block}
\end{equation}
In a $x$-\textit{oriented grid} $Q = N_{y}$, $P = N_{x}$ and $q$ and $p$ can be defined by the functions:
\begin{equation}
q(i, j) = \; \mid l(i) - l(j) \mid
\label{eq:Hxx-q-x-oriented}
\end{equation}
and
\begin{equation}
p(i, j) = \; \mid k(i) - k(j) \mid \quad ,
\label{eq:Hxx-p-x-oriented}
\end{equation}
where $l(i)$ and $l(j)$ are defined by equation \ref{eq:l-x-oriented} 
and $k(i)$ and $k(j)$ are defined by equation \ref{eq:k-x-oriented}.
For $y$-oriented grids, $Q = N_{x}$, $P = N_{y}$ and the block indices
$q$ and $p$ are defined, respectively, by the following integer functions 
of the matrix indices $i$ and $j$:
\begin{equation}
q(i, j) = \; \mid k(i) - k(j) \mid 
\label{eq:Hxx-q-y-oriented}
\end{equation}
and
\begin{equation}
p(i, j) = \; \mid l(i) - l(j) \mid \quad ,
\label{eq:Hxx-p-y-oriented}
\end{equation}
This struture can also describe matrices $\mathbf{H_{yy}}$ and $\mathbf{H_{zz}}$ in the same manner and they are identical to the structure of the gravity sensitivity matrix from \cite{takahashi2020convolutional}.

>>>>

The computational cost required for estimating the classical solution can be very high
when dealing with large datasets. In the following subsections, we will show how to 
explore the structure of the sensitivity matrix $\mathbf{A}$ and 
efficiently solve the least-squares normal equations (equation \ref{eq:normal-equations}).

@@

To access the structure of the sensitivity matrix $\mathbf{A}$ 
(equation \ref{eq:predicted-data-vector_mag}), let us first rewrite its elements 
$a_{ij}$ (equation \ref{eq:aij_mag}) in the following way:
\begin{equation}
\begin{split}
a_{ij} = a^{xx}_{ij} + a^{xy}_{ij} + a^{xz}_{ij} + a^{yy}_{ij} + a^{yz}_{ij} + a^{zz}_{ij} \: ,
\end{split}
\label{eq:aij_mag_expand}
\end{equation}
where
\begin{equation}
a^{\alpha\beta}_{ij} = 
\begin{cases}
c_{m} \, \frac{\mu_{0}}{4\pi} 
\left( F_{\alpha} u_{\beta} \right) h^{\alpha\beta}_{ij} \: &, \quad \alpha = \beta \\
c_{m} \, \frac{\mu_{0}}{4\pi} 
\left( F_{\alpha} u_{\beta} + F_{\beta} u_{\alpha} \right) h^{\alpha\beta}_{ij} \: &, \quad \alpha \ne \beta \\
\end{cases}
\: , \quad \alpha, \beta = x, y, z \: ,
\label{eq:aij_alpha_beta}
\end{equation}
are defined by the elements of $\hat{\mathbf{F}}$ 
(equation \ref{eq:unit_vector_F}), $\hat{\mathbf{u}}$ (equation \ref{eq:u_hat}) and 
$\mathbf{H}_{ij}$ (equations \ref{eq:Hij} and \ref{eq:hij_alpha_beta}).
Then, we can rewrite the sensitivity matrix $\mathbf{A}$ 
(equation \ref{eq:predicted-data-vector_mag}) according to:
\begin{equation}
\mathbf{A} = \mathbf{A_{xx}} + \mathbf{A_{xy}} + \mathbf{A_{xz}} + 
\mathbf{A_{yy}} + \mathbf{A_{yz}} + \mathbf{A_{zz}} \: ,
\label{eq:A_expand}
\end{equation}
where $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ are $N \times M$ matrices with elements 
$ij$ defined by $a^{\alpha\beta}_{ij}$ (equation \ref{eq:aij_alpha_beta}).

@@

By using equations \ref{eq:xi}--\ref{eq:l-y-oriented} to define the coordinates $x_{i}$ and 
$y_{i}$ of the observation points and $x_{j}$ and $y_{j}$ of the equivalent sources, we can
rewrite the elements $h^{\alpha\beta}_{ij}$ (equation \ref{eq:hij_alpha_beta}) of matrix 
$\mathbf{H}_{ij}$ (equation \ref{eq:Hij}) as follows:
\begin{equation}
h^{xx}_{ij} = 
\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
\label{eq:hxx_regular}
\end{equation}
\begin{equation}
h^{yy}_{ij} = 
\frac{3 \left( \Delta l_{ij} \, \Delta_{y} \right)^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
\label{eq:hyy_regular}
\end{equation}
\begin{equation}
h^{zz}_{ij} = 
\frac{3 \Delta_{z}^{2}}{r_{ij}^{5}} - \frac{1}{r_{ij}^{3}} \: ,
\label{eq:hzz_regular}
\end{equation}
\begin{equation}
h^{xy}_{ij} = 
\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right)\left( \Delta l_{ij} \, \Delta_{y} \right)}{r_{ij}^{5}} \: ,
\label{eq:hxy_regular}
\end{equation}
\begin{equation}
h^{xz}_{ij} = 
\frac{3 \left( \Delta k_{ij} \, \Delta_{x} \right) \Delta_{z}}{r_{ij}^{5}}
\label{eq:hxz_regular}
\end{equation}
and
\begin{equation}
h^{yz}_{ij} = 
\frac{3 \left( \Delta l_{ij} \, \Delta_{y} \right) \Delta_{z}}{r_{ij}^{5}} \: ,
\label{eq:hyz_regular}
\end{equation}
where $\Delta_{z} = z_{c} - z_{0}$, 
\begin{equation}
\Delta k_{ij} = \frac{x_{i} - x_{j}}{\Delta_{x}} = k(i) - k(j) \: ,
\label{eq:Delta_kij}
\end{equation}
\begin{equation}
\Delta l_{ij} = \frac{y_{i} - y_{j}}{\Delta_{y}} = l(i) - l(j)
\label{eq:Delta_lij}
\end{equation}
and 
\begin{equation}
\frac{1}{r_{ij}} = 
\frac{1}{\sqrt{\left( \Delta k_{ij} \, \Delta_{x} \right)^{2} + \left( \Delta l_{ij} \, \Delta_{y} \right)^{2} + \Delta_{z}^{2}}} \: .
\label{eq:1_rij_regular}
\end{equation}
Note that the integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ (equations 
\ref{eq:k-x-oriented}--\ref{eq:l-y-oriented}) defining $\Delta k_{ij}$ (equation
\ref{eq:Delta_kij}), $\Delta l_{ij}$ (equation \ref{eq:Delta_lij}) and 
$\tfrac{1}{r_{ij}}$ (equation \ref{eq:1_rij_regular}) assume different 
forms depending on the grid orientation.
Despite of that, it can be shown that
\begin{equation}
\Delta k_{ij} = - \Delta k_{ji} \: ,
\label{eq:Delta_kij_symmetry}
\end{equation}
\begin{equation}
\Delta l_{ij} = - \Delta l_{ji}
\label{eq:Delta_lij_symmetry}
\end{equation}
and 
\begin{equation}
\frac{1}{r_{ij}} = \frac{1}{r_{ji}}
\label{eq:1_rij_symmetry}
\end{equation}
for any grid orientation.

@@

By using equations \ref{eq:hxx_regular}--\ref{eq:1_rij_regular} to compute 
$a^{\alpha\beta}_{ij}$ (equation \ref{eq:aij_alpha_beta}), we can show that 
matrices $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation \ref{eq:A_expand}) assume
well-defined structures that can be conveniently
represented by using \textit{block indices} $q$ and $p$ \citep{takahashi2020convolutional}.
These indices are defined by the integer functions $\Delta k_{ij}$ and $\Delta l_{ij}$ 
(equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}), in terms of the indices $i$ 
of the observation points $(x_{i}, y_{i}, z_{0})$ and $j$ of the equivalent sources
$(x_{j}, y_{j}, z_{c})$.
For $x$-\textit{oriented grids}, $Q = N_{y}$, $P = N_{x}$ and the block indices
$q$ and $p$ are given by:
\begin{equation}
q \equiv q(i, j) = \Delta l_{ij}
\label{eq:q-x-oriented}
\end{equation}
and
\begin{equation}
p \equiv p(i, j) = \Delta k_{ij} \: ,
\label{eq:p-x-oriented}
\end{equation}
where $\Delta k_{ij}$ and $\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}) 
are defined by integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ given by equations 
\ref{eq:k-x-oriented} and \ref{eq:l-x-oriented}.
For $y$-\textit{oriented grids}, $Q = N_{x}$, $P = N_{y}$ and the block indices
$q$ and $p$ are given by:
\begin{equation}
q \equiv q(i, j) = \Delta k_{ij}
\label{eq:q-y-oriented}
\end{equation}
and
\begin{equation}
p \equiv p(i, j) = \Delta l_{ij} \: ,
\label{eq:p-y-oriented}
\end{equation}
where $\Delta k_{ij}$ and $\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}) 
are defined by integer functions $k(i)$, $k(j)$, $l(i)$ and $l(j)$ given by equations 
\ref{eq:k-y-oriented} and \ref{eq:l-y-oriented}.
Equations \ref{eq:q-x-oriented}--\ref{eq:p-y-oriented} differ from those presented by
\citet{takahashi2020convolutional} in the absence of the module.

In order to describe matrix components $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ (equation
\ref{eq:A_expand}) in terms of block indices, we have to verify that:
(i) each pair of indices $i$ and $j$ produce a pair of block indices $q$ and $p$;
(ii) each element $a^{\alpha\beta}_{ij}$, defined by indices $i$ and $j$, can be 
represented by the alternative notation $a^{\alpha\beta}_{qp}$, defined in terms of
the associated block indices $q$ and $p$;
(iii) different combinations of indices $i$ and $j$ result in the same block indices $q$ and $p$; 
(iv) elements $a^{\alpha\beta}_{ij}$ having the same associated block indices $q$ and $p$
also have the same numerical value;
(v) matrix $\mathbf{A}_{\boldsymbol{\alpha\beta}}$ is formed by a grid of $Q \times Q$ blocks
$\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$, where each block has $P \times P$ elements
$a^{\alpha\beta}_{qp}$ with the same block index $q$;
(vi) blocks $\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$ disposed along the same block diagonal 
of $\mathbf{A}_{\boldsymbol{\alpha\beta}}$ have the same block index $q$ and are equal to each other;
(vii) in each block $\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$, the elements $a^{\alpha\beta}_{qp}$
having the same block index $p$ are arranged along the same diagonal and are equal to each other.

Item (i) is a direct consequence of evaluating equations \ref{eq:q-x-oriented}--\ref{eq:p-y-oriented}
for a given pair of indices $i$ and $j$. Item (ii) can be verified by substituting
the block indices $q$ and $p$ (equations \ref{eq:q-x-oriented}--\ref{eq:p-y-oriented}) 
into the second derivatives $h^{\alpha\beta}_{ij}$ 
(equations \ref{eq:hxx_regular}--\ref{eq:hyz_regular}) and then use them to
define the elements $a^{\alpha\beta}_{ij}$ (equation \ref{eq:aij_alpha_beta}). The result
will be the elements $a^{\alpha\beta}_{qp}$. Item (iii) can be shown by first verifying that
different combinations of indices $i$ and $j$ result in the same integer functions
$\Delta k_{ij}$ and $\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}).
Then, by using equations \ref{eq:q-x-oriented}--\ref{eq:p-y-oriented}, we conclude that
different combinations of $i$ and $j$ result in the same indices $q$ and $p$.
Item (iv) can be shown by noting that, according to equations 
\ref{eq:q-x-oriented}--\ref{eq:p-y-oriented}, elements $a^{\alpha\beta}_{ij}$ having the same 
associated block indices $q$ and $p$ also have the same associated integer functions 
$\Delta k_{ij}$ and $\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}).
As a consequence, the associated second derivatives $h^{\alpha\beta}_{ij}$ 
(equations \ref{eq:hxx_regular}--\ref{eq:hyz_regular}) have the same numerical values
and elements $a^{\alpha\beta}_{ij}$ as well. Items (v)-(vii) can be shown by computing the 
block indices $q$ and $p$ (equations \ref{eq:q-x-oriented}--\ref{eq:p-y-oriented}) for a small 
data grid formed by, say, $N_{x} = 3$, $N_{y} = 2$, and generalizing the result for larger grids.

From equations \ref{eq:q-x-oriented}--\ref{eq:p-y-oriented}, we can show that
the block index $q$ assume integer values varying from $-Q+1$ to $Q-1$, while 
$p$ varies from $-P+1$ to $P-1$. Positive (negative) indices $q$ are associated with blocks
$\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$ located below (above) the main block diagonal 
of $\mathbf{A}_{\boldsymbol{\alpha\beta}}$.
Similarly, positive (negative) indices $p$ are associated with elements $a^{\alpha\beta}_{qp}$ 
located below (above) the main diagonal of $\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$.
According to this notation, $q$ defines the block $\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$
and $p$ the diagonal where the element $a^{\alpha\beta}_{qp}$ is located within this block, so that
matrices $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ 
can be written, in a general form, as a grid of $Q \times Q$ blocks
\begin{equation}
\mathbf{A}_{\boldsymbol{\alpha\beta}} = \begin{bmatrix}
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{0}   & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{-1} & \cdots          & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{-Q+1} \\
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{1}   & \ddots          & \ddots          & \vdots           \\ 
\vdots           & \ddots          & \ddots          & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{-1}   \\
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{Q-1} & \cdots          & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{1}  & \mathbf{A}_{\boldsymbol{\alpha\beta}}^{0}
\end{bmatrix}_{N \times N} \: ,
\label{eq:BTTB_A_alpha_beta}
\end{equation}
with blocks $\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q}$ given by
\begin{equation}
\mathbf{A}_{\boldsymbol{\alpha\beta}}^{q} = \begin{bmatrix}
a^{\alpha\beta}_{q0}   & a^{\alpha\beta}_{q(-1)} & \cdots  & a^{\alpha\beta}_{q(-P+1)} \\
a^{\alpha\beta}_{q1}   & \ddots     & \ddots  & \vdots       \\ 
\vdots      & \ddots     & \ddots  & a^{\alpha\beta}_{q(-1)}   \\
a^{\alpha\beta}_{q(P-1)} & \cdots     & a^{\alpha\beta}_{q1}  & a^{\alpha\beta}_{q0}
\end{bmatrix}_{P \times P} \: .
\label{eq:Aq_block}
\end{equation}
This well-defined structure (equations \ref{eq:BTTB_A_alpha_beta} and \ref{eq:Aq_block}) 
assumed by matrix components $\mathbf{A_{\boldsymbol{\alpha\beta}}}$ 
(equation \ref{eq:A_expand}) is called Block-Toeplitz Toeplitz-Block (BTTB) 
\citep[e.g., ][ p. 67]{chan-jin2007}.

@@

Equations \ref{eq:BTTB_A_alpha_beta} and \ref{eq:Aq_block} define the general BTTB
structure of all matrix components $\mathbf{A_{\boldsymbol{\alpha\beta}}}$, but 
there are some differences between them.
The integer functions $\Delta k_{ij}$ and $\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and
\ref{eq:Delta_lij}) play a crucial role in defining the individual structures of matrix components 
$\mathbf{A}_{\boldsymbol{\alpha\beta}}$ (equation \ref{eq:BTTB_A_alpha_beta}).
These integer functions may assume positive and negative values depending on the relative
position of the observation point $(x_{i}, y_{i}, z_{0})$ and the equivalent source 
$(x_{j}, y_{j}, z_{c})$. 
%It is important noting that the structure of a given matrix component 
%$\mathbf{A}_{\boldsymbol{\alpha\beta}}$ is controlled by the numerator of the second derivatives
%$h^{\alpha\beta}_{ij}$ (equations \ref{eq:hxx_regular}--\ref{eq:hyz_regular}),
%$\alpha, \beta = x, y, z$, which are defined by the integer functions $\Delta k_{ij}$ and 
%$\Delta l_{ij}$ (equations \ref{eq:Delta_kij} and \ref{eq:Delta_lij}).

Let us consider the matrix component $\mathbf{A}_{\boldsymbol{xx}}$, with elements
$a^{xx}_{ij}$ (equation \ref{eq:aij_alpha_beta}) defined by the second derivative
$h^{xx}_{ij}$ (equation \ref{eq:hxx_regular}). It can be easily verified from equations
\ref{eq:Delta_kij_symmetry} and \ref{eq:1_rij_symmetry} that $h^{xx}_{ij} = h^{xx}_{ji}$.
As a consequence, $a^{xx}_{ij} = a^{xx}_{ji}$ and $\mathbf{A}_{\boldsymbol{xx}}$ is symmetric.
Now, let us investigate the elements $a^{xx}_{qp}$ forming the blocks $\mathbf{A}_{\boldsymbol{xx}}^{q}$.
For $x$-oriented grids, the block indices $q$ and $p$ are defined by equations \ref{eq:q-x-oriented} and 
\ref{eq:p-x-oriented} and $a^{xx}_{qp}$ can be written as follows:
\begin{equation}
a^{xx}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
\left( F_{x} u_{x} \right) \frac{3 \left( p \, \Delta_{x} \right)^{2}}{r_{qp}^{5}} - 
\frac{1}{r_{qp}^{3}} \: ,
\label{eq:aqp_xx_x_oriented}
\end{equation}
where
\begin{equation}
\frac{1}{r_{qp}} = 
\frac{1}{\sqrt{\left( p \, \Delta_{x} \right)^{2} + \left( q \, \Delta_{y} \right)^{2} + \Delta_{z}^{2}}} \: .
\label{eq:1_rqp_x_oriented}
\end{equation}
For $y$-oriented grids, the block indices $q$ and $p$ are defined by equations \ref{eq:q-y-oriented} and 
\ref{eq:p-y-oriented} and $a^{xx}_{qp}$ can be written as follows:
\begin{equation}
a^{xx}_{qp} = c_{m} \, \frac{\mu_{0}}{4\pi} 
\left( F_{x} u_{x} \right) \frac{3 \left( q \, \Delta_{x} \right)^{2}}{r_{qp}^{5}} - 
\frac{1}{r_{qp}^{3}} \: ,
\label{eq:aqp_xx_y_oriented}
\end{equation}
where
\begin{equation}
\frac{1}{r_{qp}} = 
\frac{1}{\sqrt{\left( q \, \Delta_{x} \right)^{2} + \left( p \, \Delta_{y} \right)^{2} + \Delta_{z}^{2}}} \: .
\label{eq:1_rqp_y_oriented}
\end{equation}
From equations \ref{eq:aqp_xx_x_oriented}--\ref{eq:1_rqp_y_oriented}, we can easily verify that
$a^{xx}_{qp} = a^{xx}_{(-q)p}$ and that $a^{xx}_{qp} = a^{xx}_{q(-p)}$ for any grid orientation.
It means that the blocks $\mathbf{A}_{\boldsymbol{xx}}^{q} = \mathbf{A}_{\boldsymbol{xx}}^{(-q)}$
and that they are symmetric. From this results we conclude the matrix component 
$\mathbf{A}_{\boldsymbol{xx}}$ is symmetric-Block-Toeplitz symmetric-Toeplitz-Block for any 
grid orientation.
The same reasoning can be used to show that matrices $\mathbf{A}_{\boldsymbol{yy}}$ and
$\mathbf{A}_{\boldsymbol{zz}}$ also have this structure.

@@
